---
title: "2_yojoa_aquasat_where_you_at?"
author: "Matthew Ross"
date: "7/31/2019"
output: html_document
---


## Yojoa-AquaSat where you at?

In order to make predictions in Lake Yojoa (a relatively data poor lake), we
want to use AquaSat (a dataset of 600,000 matchups) to train a machine learning
algorithm. However, we need to first know, what lakes (in AquaSat) are most like 
Yojoa. Here's an attempt at answering that question

```{r include=F}

library(feather)
library(tidyverse)
library(sf)
library(mapview)
library(lubridate)
library(class)
library(dissUtils)
library(vegan)
library(BART) #bayesian regression trees, propagate uncertainty
load('data/yojoa_all.RData')

## Subset yojoa to the most representative sites
yojoa_representative <- passes %>%
  #Keep sites Casey says represent lake color generally
  filter(ID %in% c('B','F','E','P','R')) %>%
  #drop factors
  mutate(ID = as.character(ID)) %>%
  filter(pixelCount > 9) %>%
  #Drop sites with too much variation in swir (which should be low) 
  filter(swir1_sd < 50) %>%
  #remove cloudy images 
  filter(Cloud_Cover < 50) %>%
  #Add some band ratios from topp et al
  mutate(br = blue/red,
         gr= green/red,
         bg= blue/green,
         bn= blue/nir)

# Additional filtering on the yojoa matchup data
aq_yj <- all %>% 
  #Match secchi  names
  rename(secchi=Secchi) %>%
  #Drop nas
  filter(!is.na(secchi)) %>%
  #Drop dupes
  distinct(date,lat,long,.keep_all=T) %>%
  filter(pixelCount > 9) %>% 
  #Add ratios
  mutate(br = blue/red,
         gr= green/red,
         bg= blue/green,
         bn= blue/nir) %>%
  #Same optical filtering as above
  filter(swir1_sd < 50)  %>%
  ungroup()



#Load aquasat and do some filtering for nearest
#neighbor matching to yojoa. 

aq <- read_feather('data/wq_rs_join_v2.feather') %>%
  #Keep only lakes
  filter(type=='Lake') %>%
  #Drop NAs
  filter(!is.na(secchi)) %>%
  # Keep sites with lots of pixels
  filter(pixelCount > 50) %>%
  # Keep sites that are generally called water by Pekel et al., 
  filter(pwater > 90) %>%
  # Set some sensibility thresholds in optical space
  filter_at(vars(blue,green,red,nir,swir1,swir2),any_vars(.>0 & .<3000)) %>%
  # Drop cloudy images
  filter(clouds < 40) %>% 
  #Same filters as above
  filter(swir1_sd < 50) %>%
  mutate(br = blue/red,
         gr= green/red,
         bg= blue/green,
         bn= blue/nir) %>%
  #Tropical match, tropics are hot, keep summer only.
  filter(month(date) %in% c(6,7,8,9)) %>%
  #Keep data that matches distribution in yojoa
  filter(secchi > min(situ$Secchi,na.rm=T)-0.1 & secchi < max(situ$Secchi,na.rm=T)+0.1) 



```

## AquaSat similar site selection

### Match sites by satellite

Don't mess with this 
```{r}

#Nearest 3 neighbors within each satellite
matcher <- function(a=yojoa_representative,b=aq,satellite=5){
  a <- a %>%
    filter(sat == satellite) %>%
    select(red,blue,green,nir) %>%
    as.matrix()
  
  b <- aq %>% 
    filter(sat == satellite) %>%
    select(red,blue,green,nir) %>%
    as.matrix()
  
  distance <- diss(a,b)
    neighbors <- neighbors(a,b,n.neighbors=3) 
    neighbor.index <- neighbors.identify(neighbors,distance) %>% 
    as.vector()


  match <- aq[neighbor.index,] %>%
    mutate(similarity = neighbors %>% as.vector())

  return(match)
}


m5 <- matcher(satellite=5)
m7 <- matcher(satellite = 7)
m8 <- matcher(satellite = 8)



#Set a minimum similarity score
#Bind satellites
good_sites <- rbind(m5,m7,m8) %>%
  #Cluster by siteid
  group_by(SiteID) %>%
  # Get a count 
  mutate(count=n()) %>%
  #drop groups
  ungroup() %>%
  #sort by counts
  arrange(count) %>%
  filter(similarity < 250) %>% 
  filter(swir1_sd < 50)



```








### Where are the most similar sites

```{r}
train <- good_sites %>%
  filter(similarity < 20) %>%
  filter(count > 2) 

#Make a map of the matched sites
site_map <-  train %>%
  st_as_sf(.,coords=c('long','lat'),crs=4326) 

mapview(site_map,zcol='count')
```


```{r}

# TUNELANDIA. This can be played with a bunch and you should read about it
#and play. 
ag_mod <- wbart(x.train=train %>%
                  dplyr::select(blue,green,red,nir,br,gr,bg,bn,zenith) %>% as.matrix(),
                y.train = train %>% pull(secchi),
                x.test = aq_yj %>% 
                  select(blue,green,red,nir,br,gr,bg,bn,zenith) %>% as.matrix(),
                cont=T,k=2,
                ntree=200,numcut=20)

#Tune ntree, k, numcut, bands that you use, etc...

aq_yj$bpred <- ag_mod$yhat.test.mean


error <- 
  tibble(rmse=Metrics::rmse(aq_yj$bpred,aq_yj$secchi),
                     mdae=Metrics::mdae(aq_yj$bpred,aq_yj$secchi),
                     mape=Metrics::mape(aq_yj$bpred,aq_yj$secchi),
                     bias=Metrics::bias(aq_yj$bpred,aq_yj$secchi))



error

ggplot(aq_yj,aes(x=secchi,y=bpred,color=year(date))) + 
  geom_point() + 
  geom_abline(intercept=0,slope=1) + 
  stat_smooth(method='lm',color='blue')

```

## Raw yojoa vs bands

```{r}
ggplot(aq_yj,aes(x=br,y=secchi,color=month(date))) + 
  geom_point() + 
  stat_smooth(method='lm',color='blue')
```


```{r}



mcmc_predict = predict(ag_mod,yojoa_representative %>% select(blue,green,red,nir,br,
                                                          gr,bg,bn,zenith) %>%
           as.matrix())

means <- colMeans(mcmc_predict)

medians <- apply(mcmc_predict,2,median)

q95er <- function(x,y=0.95){quantile(x,y)}
q95 <- apply(mcmc_predict,2,q95er)
q5 <- apply(mcmc_predict,2,q95er,y=0.05)

e <- yojoa_representative %>%
  mutate(secchi = means,
         s95 = q95,
         s5 = q5)

gp <- ggplot(e %>%
               filter(ID == 'E'),aes(x=date,y=secchi,color=ID)) + 
  geom_line()


e.means <- e %>%
  filter(ID == 'E') %>%
  #filter(month(date) %in% c(1:5)) %>%
  mutate(years=year(date)) %>%
  group_by(years) %>%
  summarize(secchi=median(secchi),
            s95 = median(s95),
            s5 = median(s5),
            count=n()) %>%
  filter(count > 2)


# monthlies <- e %>%
#   mutate(month=month(date),
#          year=year(date)) %>%
#   group_by(month,year) %>%
#   summarize(secchi=mean(secchi))  %>%
#   ungroup()
# 
# ggplot(monthlies,aes(x=month,y=secchi,color=year,group=year)) + 
#   geom_line()

situ_means <- situ %>%
  filter(Location %in% c('Index','E','Basterrechea Index')) %>%
  mutate(years=year(date)) %>%
  group_by(years) %>%
  summarize(secchi=mean(Secchi,na.rm=T),
            count=n()) %>%
  filter(count > 2) 

str(e$datetime)
ggplot(e,aes(x=date,y=secchi)) + 
  geom_ribbon(aes(ymin=s5,ymax=s95),fill='gray',color=NA,alpha=0.5) + 
  geom_point() +
  geom_point(data=situ,aes(x=situ_date,y=Secchi),color='red')


```

## Yojoa self train

Just for your comparison
```{r}
aq_yj$index <- 1:nrow(aq_yj)

aq_train <- aq_yj %>%
  sample_frac(0.5)

aq_test <- aq_yj %>%
  filter(!index %in% aq_train$index)


ag_mod <- wbart(x.train=aq_train %>%
                  dplyr::select(blue,green,red,nir,br,gr,bg,bn,zenith) %>% as.matrix(),
                y.train = aq_train %>% pull(secchi),
                x.test = aq_test %>% 
                  select(blue,green,red,nir,br,gr,bg,bn,zenith) %>% as.matrix(),
                cont=T,k=2,
                ntree=200,numcut=20)

aq_test$bpred <- ag_mod$yhat.test.mean


error <- 
  tibble(rmse=Metrics::rmse(aq_test$bpred,aq_test$secchi),
                     mdae=Metrics::mdae(aq_test$bpred,aq_test$secchi),
                     mape=Metrics::mape(aq_test$bpred,aq_test$secchi),
                     bias=Metrics::bias(aq_test$bpred,aq_test$secchi))


error

ggplot(aq_test,aes(x=secchi,y=bpred,color=month(date))) + 
  geom_point() + 
  geom_abline(intercept=0,slope=1) + 
  stat_smooth(method='lm',color='blue') + 
  ggtitle('selftrain on yojoa only')



```
